Non-word repetition in Yélî Dnye

### 2020-09-14 manuscript.tex
addressed middys comments code review

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  consent, ``correct'', ``accuracy'', ``performance''
\item
  draft discussion
\end{itemize}

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood (e.g., Hazan \& Barrett, 2000). One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. Accuracy is thought to reflect long-term phonological knowledge (which allows the child to perceive the item accurately even though it is not a real word they have encountered before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item accurately even though it had never been pronounced before).

@@

In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. The present report uses these unique data to contribute to several lines of research. First, we made sure that some of the non-word items contained typologically rare and/or challenging sounds, so that we can contribute to the study of whether rare sounds are disadvantaged in perception and/or production, both in terms of overall accuracy of repetition and patterns of mispronunciation. Second, we look at the effects of item length, since previous NWR research has uncovered variable effects. Third, we contribute to the basic NWR literature by adding a data point for children learning a rarely-studied language and culture, and further investigating whether there are structured sources that account for individual variation.

@@

Previous work seems to avoid difficult sounds, but we felt this was important to represent Yélî Dnye, so we also varied this factor. We designed a relatively large number of items but, aware that this may render the task longer and more tiresome, we split some of the items across children. This allowed us to get information about repetition accuracy of more items.

@@

\caption{\label{tab:tab2-stims}NWR stimuli in orthographic and phonological representations.}
\centering
\begin{tabular}[t]{llllllllll}
\toprule
\multicolumn{2}{c}{Practice} & \multicolumn{2}{c}{Monosyll} & \multicolumn{2}{c}{Bisyll} & \multicolumn{2}{c}{Trisyll} & \multicolumn{2}{c}{Tetrasyll} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon.\\
\midrule
nopimade & nɔpimæṭɛ & dp:a & ṭp̃æ & kamo & kæmɔ & dimope & ṭimɔpɛ & dipońate & ṭipɔnætɛ\\
poni & pɔni & dpa & ṭpæ & kańi & kæni & diyeto & ṭijɛtɔ & ńomiwake & nɔmiwækɛ\\
wî & wɯ & dpâ & ṭpɑ & kipo & kipɔ & meyadi & mɛjæṭi & todiwuma & tɔṭiwumæ\\
 &  & dpê & ṭpə & ńoki & nɔki & mituye & mitujɛ & wadikeńo & wæṭikɛnɔ\\
 &  & dpéé & ṭpeː & ńomi & nɔmi & ńademo & næṭɛmɔ &  & \\
\addlinespace
 &  & dpi & ṭpi & piwa & piwæ & ńayeki & næjɛki &  & \\
 &  & dpu & ṭpu & towi & tɔwi & ńuyedi & nujɛṭi &  & \\
 &  & gh:ââ & ɣ̃ɑː & tupa & tupæ & pedumi & pɛṭumi &  & \\
 &  & ghuu & ɣuː &  &  & tiwuńe & tiwunɛ &  & \\
 &  & kpu & kpu &  &  & tumowe & tumɔwɛ &  & \\
\addlinespace
 &  & lv:ê & lβʲ̃ə &  &  & widońe & wiṭɔnɛ &  & \\
 &  & lva & lβʲæ &  &  & wumipo & wumipɔ &  & \\
 &  & lvi & lβʲi &  &  &  &  &  & \\
 &  & tp:a & tp̃æ &  &  &  &  &  & \\
 &  & tpâ & tpɑ &  &  &  &  &  & \\
\addlinespace
 &  & tpê & tpə &  &  &  &  &  & \\
\bottomrule
\end{tabular}
\end{table}

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level accuracy (e.g., Cristia et al., 2020). We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being ``bilik'', and imagine a response ``bilig''. This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80\% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20\% (20\% of phonemes were substituted or deleted).\\

@@

This study was approved as part of a larger research effort by Marisa Casillas, with the research protocol having been approved by the IRB committee entitled The Ethics Committee of the Faculty of Social Sciences (Ethiek Commissie van de faculteit der Sociale Wetenschappen; ECSW) of Radboud University, Nijmegen, The Netherlands. The approval for work outside of the Netherlands came as an amendment to the general research line on language development research (original request: ECSW2017-3001-474 Manko-Rowland; amendment: ECSW-2018-041). Participation was voluntary, with children being invited to come and participate. Regardless of how they performed, children were provided with a snack as compensation. Children who came up to participate but then refused were nonetheless provided with the snack.

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = 39, SD = 26)
were on average less accurate than first ones (M = 65, SD = 15), t(38) = 6.62, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

The overall NWR score was M = 61\% (SD = 15\%). Scores based on phonemes are even higher M = 78\% (SD = 9\%). The phoneme-based normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

In this analysis, we were interested in variation in performance as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID \& target ID as random variables.

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.77, SE ß = 0.19, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Age was also a significant predictor (ß = 0.31, SE ß = 0.13, p = 0.02).

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings ({\textbf{???}}), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(23)=0.65, =0). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.50, SE ß = 0.25, p = 0.04) and age (ß = 0.31, SE ß = 0.13, p = 0.02) were significant predictors of accuracy, but corpus frequency (ß = 0.37, SE ß = 0.24, p = 0.12) was not.

@@

\textbf{Option 1: just numbers, first attempts only }
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 7 vowels deleted, and 3 consonants.

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels, 0 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (43 oral vowels produced with a different quality; 0 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0) than vice versa (0). Simple consonants were mispronounced as other simple consonants quite frequently (73 simple consonants mispronounced as other simple ones, compared to 0 complex ones).

\textbf{Option 2: numbers and proportions, first attempts only}
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 7 vowels deleted (about 0.54\% of all vowel targets), and 3 consonants deleted (about 0.17\% of all consonant targets).

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels or about 0\% out of all oral vowel targets, 0 nasal target vowels produced as oral vowels or about NA\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (43 oral vowels produced with a different quality or about 0.29\% out of all oral vowel targets; 0 nasal vowels produced with a different quality or about NA\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0 times or about NA\% out of all complex consonant targets) than vice versa (0 times or about NA\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (73 simple consonants mispronounced as other simple ones or about 0.35\% out of all simple consonant targets, compared to 0 complex ones or about NA\% out of all complex consonant targets).

\textbf{Option 3: numbers and proportions, ALL attempts}
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 11 vowels deleted (about 0.25\% of all vowel targets), and 8 consonants deleted (about 0.21\% of all consonant targets).

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (52 oral target vowels produced as nasal vowels or about 2.01\% out of all oral vowel targets, 58 nasal target vowels produced as oral vowels or about 6.46\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (205 oral vowels produced with a different quality or about 3.22\% out of all oral vowel targets; 23 nasal vowels produced with a different quality or about 7.08\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 250 times or about 5.57\% out of all complex consonant targets) than vice versa (2 times about 5.57\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (129 simple consonants mispronounced as other simple ones or about 1.14\% out of all simple consonant targets, compared to 55 complex ones or about 2.76\% out of all complex consonant targets).

@@

\hypertarget{accuracy-as-a-function-of-item-length}{%
\subsubsection{Accuracy as a function of item length}\label{accuracy-as-a-function-of-item-length}}

Next, we inspected whether accuracy varied as a function of word length. Results are shown on table XX. This table shows that monosyllables accuracy was much lower than other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of decreased accuracy for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.

@@

\caption{\label{tab:tablength}NWR scores measured in whole word accuracy, phoneme-based accuracy, and normalized Levenshtein Distance, separately for the four stimuli lengths.}

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 59 (17) & 41 (17)\\
79 (22) & 94 (7) & 6 (7)\\
78 (19) & 94 (6) & 6 (6)\\
75 (32) & 93 (11) & 7 (11)\\
\bottomrule
\end{tabular}

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,548.04) = 0.33, p = 0.12). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.18) = 21.34, p = 0.86), birth order (data missing for 17 children, 2,678.47) = -0.16, p = 0.44) or maternal education (data missing for 2 children, 2,091.92) = 0.09, p = 0.67)

@@

\begin{itemize}
\item
  What is the overall repetition accuracy (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllables, sound complexity)?
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Children are more accurate for mono-syllables than longer items
\item
  The length distribution in Yélî words is more balanced than that in English, and thus the performance decline for poly- versus mono-syllables may be less pronounced than that for English. \textbf{Check for work on European languages that may have looked into this}
\item
  Similarly, we do not know of NWR research that manipulates the difficulty of the sounds that are included in the items, but word naming and other research suggests that children are more accurate when producing easy and/or typologically common sounds than difficult and/or typologically rare sounds {[}CITE{]}. Therefore, we expect higher accuracy for items with common sounds than in those with rare sounds.
\item
  The Yélî sound inventory is very large and compressed, with many similar sounds that are acoustic and articulatory neighbors. Therefore, this may constitute a pressure for children to have finer auditory skills (and perhaps more precise articulations) than children speaking languages with a simpler inventory. As a result, differences between easier and harder items may be smaller in this work than in other research. \textbf{no work looking at consonants \& vowels? no work looking at nasal vowels in particular?}
\end{enumerate}

>>>>

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. NWR scores are thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated without changes less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  draft discussion
\end{itemize}

\hypertarget{todo-alex}{%
\subsection{TODO Alex}\label{todo-alex}}

\begin{itemize}
\tightlist
\item
  add effect sizes to all analyses
\item
  reply to questions
\end{itemize}

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood (e.g., Hazan \& Barrett, 2000). One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. NWR scores are thought to reflect long-term phonological knowledge (which allows the child to perceive the item precisely even though it is not a real word they have encountered before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely even though it had never been pronounced before).

@@

In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. The present report uses these unique data to contribute to several lines of research. First, we made sure that some of the non-word items contained typologically rare and/or challenging sounds, so that we can contribute to the study of whether rare sounds are disadvantaged in perception and/or production, both in terms of overall repetition scores and patterns of mispronunciation. Second, we look at the effects of item length, since previous NWR research has uncovered variable effects. Third, we contribute to the basic NWR literature by adding a data point for children learning a rarely-studied language and culture, and further investigating whether there are structured sources that account for individual variation.

@@

Previous work seems to avoid difficult sounds, but we felt this was important to represent Yélî Dnye, so we also varied this factor. We designed a relatively large number of items but, aware that this may render the task longer and more tiresome, we split some of the items across children. This allowed us to get NWR scores for more items.

@@

\caption{\label{tab:tab2-stims}NWR stimuli in orthographic and phonological representations.}
\centering
\begin{tabular}[t]{llllllllll}
\toprule
\multicolumn{2}{c}{Practice} & \multicolumn{2}{c}{Monosyll} & \multicolumn{2}{c}{Bisyll} & \multicolumn{2}{c}{Trisyll} & \multicolumn{2}{c}{Tetrasyll} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon. & Orth. & Phon.\\
\midrule
nopimade & nɔpimæṭɛ & dp:a & ṭp̃æ & kamo & kæmɔ & dimope & ṭimɔpɛ & dipońate & ṭipɔnætɛ\\
poni & pɔni & dpa & ṭpæ & kańi & kæni & diyeto & ṭijɛtɔ & ńomiwake & nɔmiwækɛ\\
wî & wɯ & dpâ & ṭpɑ & kipo & kipɔ & meyadi & mɛjæṭi & todiwuma & tɔṭiwumæ\\
 &  & dpê & ṭpə & ńoki & nɔki & mituye & mitujɛ & wadikeńo & wæṭikɛnɔ\\
 &  & dpéé & ṭpeː & ńomi & nɔmi & ńademo & næṭɛmɔ &  & \\
\addlinespace
 &  & dpi & ṭpi & piwa & piwæ & ńayeki & næjɛki &  & \\
 &  & dpu & ṭpu & towi & tɔwi & ńuyedi & nujɛṭi &  & \\
 &  & gh:ââ & ɣ̃ɑː & tupa & tupæ & pedumi & pɛṭumi &  & \\
 &  & ghuu & ɣuː &  &  & tiwuńe & tiwunɛ &  & \\
 &  & kp:ââ & kp̃ɑː &  &  & tumowe & tumɔwɛ &  & \\
\addlinespace
 &  & kpu & kpu &  &  & widońe & wiṭɔnɛ &  & \\
 &  & lv:ê & lβʲ̃ə &  &  & wumipo & wumipɔ &  & \\
 &  & lva & lβʲæ &  &  &  &  &  & \\
 &  & lvi & lβʲi &  &  &  &  &  & \\
 &  & tp:a & tp̃æ &  &  &  &  &  & \\
\addlinespace
 &  & tpâ & tpɑ &  &  &  &  &  & \\
 &  & tpê & tpə &  &  &  &  &  & \\
\bottomrule
\end{tabular}
\end{table}

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring (e.g., Cristia et al., 2020). We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being ``bilik'', and imagine a response ``bilig''. This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80\% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20\% (20\% of phonemes were substituted or deleted).\\

@@

This study was approved as part of a larger research effort by Marisa Casillas, with the research protocol having been approved by the IRB committee entitled The Ethics Committee of the Faculty of Social Sciences (Ethiek Commissie van de faculteit der Sociale Wetenschappen; ECSW) of Radboud University, Nijmegen, The Netherlands. The approval for work outside of the Netherlands came as an amendment to the general research line on language development research (original request: ECSW2017-3001-474 Manko-Rowland; amendment: ECSW-2018-041). Participation was voluntary, with children being invited to come and participate. Regardless of whether they did any of the task, children were provided with a snack as compensation. Children who came up to participate but then refused were nonetheless provided with the snack.

@@

We first checked whether NWR scores varied between first and subsequent presentations of an item by averaging word-level scores at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level scores became more heterogeneous in subsequent repetitions. Surprisingly, NWR scores for subsequent repetitions (M = 40, SD = 26)
were on average lower than first ones (M = 64, SD = 15), t(38) = 6.30, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since scores declined and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

The overall NWR score was M = 61\% (SD = 15\%). Scores based on phonemes are even higher M = 100\% (SD = 0\%). The phoneme-based normalized Levenshtein distance was M = 0\% (SD = 0\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

In this analysis, we were interested in variation in NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID \& target ID as random variables.

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.79, SE ß = 0.19, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Age was also a significant predictor (ß = 0.33, SE ß = 0.13, p = 0.01).

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings ({\textbf{???}}), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(23)=0.65, =0). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.51, SE ß = 0.25, p = 0.04) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of word-level NWR scores, but corpus frequency (ß = 0.39, SE ß = 0.24, p = 0.11) was not.

@@

\textbf{Option 1: just numbers, first attempts only }
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 0 vowels deleted, and 0 consonants.

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels, 0 nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (0 oral vowels produced with a different quality; 0 nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0) than vice versa (0). Simple consonants were mispronounced as other simple consonants quite frequently (0 simple consonants mispronounced as other simple ones, compared to 0 complex ones).

\textbf{Option 2: numbers and proportions, first attempts only}
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 0 vowels deleted (about NA\% of all vowel targets), and 0 consonants deleted (about NA\% of all consonant targets).

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels or about NA\% out of all oral vowel targets, 0 nasal target vowels produced as oral vowels or about NA\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (0 oral vowels produced with a different quality or about NA\% out of all oral vowel targets; 0 nasal vowels produced with a different quality or about NA\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0 times or about NA\% out of all complex consonant targets) than vice versa (0 times or about NA\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (0 simple consonants mispronounced as other simple ones or about NA\% out of all simple consonant targets, compared to 0 complex ones or about NA\% out of all complex consonant targets).

\textbf{Option 3: numbers and proportions, ALL attempts}
Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 0 vowels deleted (about NA\% of all vowel targets), and 0 consonants deleted (about NA\% of all consonant targets).

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (0 oral target vowels produced as nasal vowels or about NA\% out of all oral vowel targets, 0 nasal target vowels produced as oral vowels or about NA\% out of all nasal vowel targets). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (0 oral vowels produced with a different quality or about NA\% out of all oral vowel targets; 0 nasal vowels produced with a different quality or about NA\% out of all nasal vowel targets). As for consonants, asymmetries were very marked with more complex consonants (specifically dptpkpkmknmbghlv) mispronounced as simple ones (specifically mnlwyvdgptkfhch, 0 times or about NA\% out of all complex consonant targets) than vice versa (0 times about NA\% out of all simple consonant targets). Simple consonants were mispronounced as other simple consonants quite frequently (0 simple consonants mispronounced as other simple ones or about NA\% out of all simple consonant targets, compared to 0 complex ones or about NA\% out of all complex consonant targets).

@@

\hypertarget{nwr-scores-as-a-function-of-item-length}{%
\subsubsection{NWR scores as a function of item length}\label{nwr-scores-as-a-function-of-item-length}}

Next, we inspected whether NWR scores varied as a function of word length. Results are shown on table XX. This table shows that scores for monosyllables was much lower than for other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.

@@

\caption{\label{tab:tablength}NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.}

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 100 (0) & 0 (0)\\
79 (22) & 100 (0) & 0 (0)\\
78 (19) & 100 (0) & 0 (0)\\
75 (32) & 100 (0) & 0 (0)\\
\bottomrule
\end{tabular}

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,485.88) = 0.35, p = 0.09). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.08) = 21.39, p = 0.94), birth order (data missing for 17 children, 2,730.78) = -0.19, p = 0.38) or maternal education (data missing for 2 children, 2,147.84) = 0.07, p = 0.76)

@@

\begin{itemize}
\item
  What is the overall repetition scores (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllables, sound complexity)?
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Prediction based on previous work: Children have higher scores for shorter items
\item
  Prediction for Yélî made before seeing the data: The length distribution in Yélî words is more balanced than that in English, and thus the score decline for poly- versus mono-syllables may be less pronounced than that for English.
\item
  it turns out we were right! the scores for 2-4 syll words decline only slightly with length
\item
  Prediction based on previous work: Similarly, we do not know of NWR research that manipulates the difficulty of the sounds that are included in the items, but word naming and other research suggests that children achieve higher scores when producing easy and/or typologically common sounds than difficult and/or typologically rare sounds {[}CITE{]}. Therefore, we expect higher scores for items with common sounds than in those with rare sounds.
\item
  Prediction for Yélî made before seeing the data: The Yélî sound inventory is very large and compressed, with many similar sounds that are acoustic and articulatory neighbors. Therefore, this may constitute a pressure for children to have finer auditory skills (and perhaps more precise articulations) than children speaking languages with a simpler inventory. As a result, differences between easier and harder items may be smaller in this work than in other research.
\item
  it turns out we do see
\end{enumerate}
